Unifying the communicable disease spreading paradigm with Gompertzian growth

### 2022-07-18 elsarticle/elsarticle.tex
update intro

This phenomenon is recorded by others \citep{Ohnishi2020,Rypdal2020,Catala2020,rodrigues2020monte,Levitt2020}, showing Gompertz curves at national levels instead of the traditionally predicted logistic curves, and are made possible in part due to the increased cadence and quality of data acquisition compared with earlier epidemics. What is causing such a discrepancy between reality and the current theory of communicable diseases \citep{castro2020turning}?

@@

The SIRD model above also considers average macroscopic behavior of an ensemble of microscopic units justified through mean-field theory \citep{smilkov2014beyond} which does not consider network effects explicitly. Rather, all entities are connected and communicating instantaneously as shown by Mombach et al. \cite{mombach2002mean}.
However, even with such strong assumptions, it is odd that observed mortality never exhibits a stage of near-exponential growth as this macroscopic SIRD model predicts, but rather a constant negative slope in the double-log domain.

Other more general observations that conflict with these communicable disease models is 1) the lack of correlation between population density and mortality (or infection) rates \citep{Hamidi2020,Hamidi2020a,Carozzi2020,khavarian2021high,barak2021urban}, 2) no clear association between intergenerational relationships and fatality rates \citep{Arpino2020}, and 3) the widespread morbidity and mortality spikes seen during spring 2020 among birds \citep{Fischer2021}, dogs \citep{PeriseBarrios2021}, rabbits \citep{Duff2020,Hu2020,Fukui2021}, elephants \citep{Aarde2021}, and horses \citep{Kambayashi2021,King2020,Lu2020,CastilloOlivares2020}.
We are thus prompted to look for another model which can explain the observed mortality patterns.

>>>>

This phenomenon is recorded by others \citep{Ohnishi2020,Rypdal2020,Catala2020,rodrigues2020monte,Levitt2020}, showing Gompertz curves at national levels instead of the traditionally predicted logistic curves, and are made possible in part due to the increased cadence and quality of data acquisition compared with earlier epidemics. Other observations that conflict more generally with the communicable disease models are 1) the lack of correlation between population density and mortality (or infection) rates \citep{Hamidi2020,Hamidi2020a,Carozzi2020,khavarian2021high,barak2021urban}, 2) no clear association between inter-generational relationships and fatality rates \citep{Arpino2020}, and 3) the widespread morbidity and mortality spikes seen during spring 2020 among birds \citep{Fischer2021}, dogs \citep{PeriseBarrios2021}, rabbits \citep{Duff2020,Hu2020,Fukui2021}, elephants \citep{Aarde2021}, and horses \citep{Kambayashi2021,King2020,Lu2020,CastilloOlivares2020}.

What is causing such a discrepancy between these observations and the current theory of communicable diseases \citep{castro2020turning}?

@@

The SIRD model above also considers average macroscopic behavior of an ensemble of microscopic units justified through mean-field theory \citep{smilkov2014beyond} which does not consider network effects explicitly. Rather, all entities are connected and communicating instantaneously as shown by Mombach et al. \cite{mombach2002mean}.
However, even with such strong assumptions, it is odd that observed mortality never exhibits a stage of near-exponential growth as this macroscopic SIRD model predicts, but rather a constant negative slope in the double-log domain. We are thus prompted to look for another model which can explain the observed mortality patterns.
